Malacology: A Programmable Storage System Built on Ceph

### 2016-05-10 paper/paper.tex
commit comment
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

In this paper we investigate an approach that focusses on generalizing
existing storage system resources, services, and abstractions that in
generalized form can be used to \emph{program} new services. By doing so
one can reuse subsystems and their optimizations and leverage their
established correctness, robustness, and efficiency. We will refer to
this programmability as \emph{programmable storage}, which differs from
\emph{active storage} (the injection and execution of arbitrary code in
a storage system) and \emph{software-defined storage} (the control of
thin-provisioning of storage).

@@

To illustrate the benefits and challenges of this approach we have
designed and evaluated Malacology, a programmable storage system for
constructing new functionality by re-purposing existing subsystem
abstractions contained in Ceph. We build the framework on Ceph by
leveraging a broad spectrum of existing services, including distributed
locking and caching services provides by MDSs, durability and logical
storage devices provided by RADOS, and propogation of consistent cluster
state provided by the monitoring service (see \ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
functionality necessary for implementing new services. Our contributions
are:

@@

The second approach that applications may use to adapt to a storage system
limitation is to
extend the responsibility of the application. This means changing the
application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O
parallelism in a distributed storage system.

For example, SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific data (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. %This is not a
%bad proposition, but 
Changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

@@

When both the bolt-on approach and changing applications fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system. The
difficulty of both of these approaches has given rise to a third
technique called active storage.

@@

The object interface facility provided by Ceph is more than
a proof-of-concept active storage implementation.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

The popularity of the active storage component of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default
algorithms/tunables of the storage system are insufficient for the
application's performance goals, (2) programmers are becoming more aware
of their application's behavior, and (3) programmers know how to manage
resources to improve performance. Programmers gravitate towards object
interfaces because it gives them the ability to tell the storage system
about their application: if it is CPU or IO bound, if it has locality,
if its size has the potential to overload a single proxy node, etc. The
programmers know what the problem is and how to solve it, but until
the ability to modify object interfaces, they had no way to express to the storage system how to handle
their data.

@@

The Ceph team wants to merge Mantle because the scriptability is useful
for debugging, controlling the metadata balancer, and examining
trade-offs for different balancers. Unfortunately, this research quality
system is not as robust as Ceph and the Ceph team wants more safety,
durability, and consistency for the new functionality. For example, the
original Mantle API is fragile: the API is not enforced by the runtime,
the system allows injectable strings through the admin daemon,
constructing the Lua script in C++ is clunky, and the administrator can
inject really bad policies (e.g., while 1) that brings the whole system
down.

>>>>

In this paper we investigate an approach that focusses on generalizing
existing storage system resources, services, and abstractions that in
generalized form can be used to \emph{program} new services. By doing so
one can reuse subsystems and leverage their optimizations,
established correctness, robustness, and efficiency. We will refer to
this programmability as \emph{programmable storage}, which differs from
\emph{active storage} (the injection and execution of arbitrary code in
a storage system) and \emph{software-defined storage} (the control of
thin-provisioning of storage).

@@

To illustrate the benefits and challenges of this approach we have
designed and evaluated Malacology, a programmable storage system for
constructing new services by re-purposing existing subsystem
abstractions contained in Ceph. We build the framework in Ceph by
leveraging a broad spectrum of existing services, including distributed
locking and caching services provided by MDSs, durability and logical
storage devices provided by RADOS, and propogation of consistent cluster
state provided by the monitoring service (see \ref{fig:overview}). As we will show in this paper, this framework is expressive enough to provide the
functionality necessary for implementing new services. Our contributions
are:

@@

The second approach that applications may use to adapt to a storage system
limitation is to extend the responsibility of the application. This means
changing the application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O parallelism in a
distributed storage system.
In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
application and the Hadoop framework itself are changed to leverage the structure of
scientific data (3D array-based data, more specifically) to
improve locality, performance, and increase the number of early results. %This is not a
%bad proposition, but 
However, changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

@@

When both the bolt-on approach and changing applications fail to meet the needs of the application,
developers turn their attention to the storage system itself.
Traditionally, storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system.
%The
%difficulty of both of these approaches has given rise to a third
%technique called active storage.

@@

The object interface facility provided by Ceph is more than
a proof-of-concept implementation resembling active storage.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

The popularity of the custom object interface facility of Ceph hint at three
trends in the Ceph community: (1) increasingly, the default
algorithms/tunables of the storage system are insufficient for the
application's performance goals, (2) programmers are becoming more aware
of their application's behavior, and (3) programmers know how to manage
resources to improve performance. Programmers gravitate towards object
interfaces because it gives them the ability to tell the storage system
about their application: if it is CPU or IO bound, if it has locality,
if its size has the potential to overload a single proxy node, etc. The
programmers know what the problem is and how to solve it, but until
the ability to modify object interfaces, they had no way to express to the storage system how to handle
their data.

@@

The Ceph team wants to merge Mantle because the scriptability is useful
for debugging, controlling the metadata balancer, and examining
trade-offs for different balancers. Unfortunately, this research quality
system is not as robust as Ceph and the Ceph team wants more safety,
durability, and consistency for the new services. For example, the
original Mantle API is fragile: the API is not enforced by the runtime,
the system allows injectable strings through the admin daemon,
constructing the Lua script in C++ is clunky, and the administrator can
inject really bad policies (e.g., while 1) that brings the whole system
down.
